---
title: "Lego Analysis"
format: 
  revealjs:
    theme: sky
    slideNumber: true
    fontSize: 18
    transition: fade
editor: visual
---

```{r}
#| label: Loading libraries
#| echo: false
#| message: false
#| warning: false

# Loading libraries
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(scales)
library(kableExtra)
```

```{r}
#| label: Loading data
#| echo: false
#| message: false
#| warning: false

# Read data from tidytuesday
tuesdata = tidytuesdayR::tt_load('2022-09-06')

for (name in names(tuesdata)) 
{var_name = gsub("\\.csv$", "", name) 
  assign(var_name, tuesdata[[name]])}

# Dropping data not relevant for analysis
rm(tuesdata, inventory_sets, 
   elements, part_categories, 
   part_relationships, minifigs, 
   inventory_minifigs, name, var_name)
```

```{r}
#| label: Set themes
#| echo: false
#| message: false
#| warning: false

# Joining sets with themes and filtering out the sets with no parts due to error in data collection
set_themes = sets |>
  filter(!is.na(num_parts) & num_parts >0) |>
  left_join(themes |>
              select(id, theme_name = name), 
              by = c(theme_id = "id"))
```

```{r}
#| label: fig-most_popular
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Most popular themes

# Finding out the most popular themes
most_popular = set_themes |>
  group_by(theme_name) |>
  summarize(total_sets = n(),
            median_parts = round(median(num_parts, na.rm = TRUE)),
            .groups = "drop") |>
  arrange(desc(total_sets)) |>
  head(20)

# Plotting the most popular themes
most_popular_plot = 
  ggplot(most_popular,
         aes(x = total_sets, 
             y = reorder(theme_name, total_sets))) +
  geom_bar(stat = "identity", aes(fill = median_parts)) +
  geom_text(aes(label = median_parts, 
                x = total_sets + 5),
            hjust = 0, size = 3.2) +
  scale_fill_gradient(low = "#00648C", high = "#00BC8C", limits = c(0, 2600)) +
  labs(fill = "Median parts",
       title = "All time most popular lego themes with median parts",
       x = "Total number of sets",
       y = "") +
  theme_minimal() +
  theme(plot.title.position = "plot",plot.title = element_text(hjust = 0))
```

```{r}
#| label: fig-least_popular
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Least popular themes

# Finding out the least popular themes
least_popular = set_themes |>
  group_by(theme_name) |>
  summarize(total_sets = n(),
            median_parts = round(median(num_parts, na.rm = TRUE)),
            .groups = "drop") |>
  arrange(desc(total_sets)) |>
  tail(20)

# Plotting the least popular themes
least_popular_plot = 
  ggplot(least_popular, 
         aes(x = total_sets, 
             y = reorder(theme_name, total_sets))) +
  geom_bar(stat = "identity", aes(fill = median_parts)) +
  geom_text(aes(label = median_parts),
            hjust = 0, size = 3.2) +
  scale_fill_gradient(low = "#00648C", high = "#00BC8C", limits = c(0, 2600)) +
  labs(fill = "Median parts",
       title = "All time least popular lego themes with median parts",
       x = "Total number of sets",
       y = "")+ 
  xlim(0,4)+
  theme_minimal() +
  theme(plot.title.position = "plot",plot.title = element_text(hjust = 0))
```

```{r}
#| label: fig-theme_by_parts
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Theme by parts

# Table with theme name ordered by desc of sets and their median parts
theme_name_summary = set_themes |> 
  group_by(theme_name) |>
  summarize(n_sets = n(), 
            median_parts = round(median(num_parts))) |>
  arrange(desc(n_sets)) |> 
  filter(n_sets >= 100) |>
  slice_head(n = 10)


theme_summary = knitr::kable(theme_name_summary, caption = "Theme ordered by sets with their median parts", row.names = FALSE)
```

```{r}
#| message: false
#| warning: false
#| label: fig-parts_trend
#| fig-cap: Median parts by year
#| echo: false

# Creating dataframe for part_year
part_year = set_themes |>
  group_by(year) |>
  summarise(median_parts = median(num_parts, na.rm = TRUE), .groups = "drop")

# Plotting the Line chart of median parts per set over the years
part_year_plot = ggplot(part_year, aes(x = year, y = median_parts)) +
  geom_area(fill = "#00BC8C", alpha = 0.2) +
  geom_line(color = "#00BC8C", linewidth = 1) +
  geom_point(color = "#00648C") +
  geom_smooth(method = "loess", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Median number of parts over the years",
       x = "Year",
       y = "Median parts per set")+
  theme_minimal() 
```

```{r}
#| message: false
#| warning: false
#| label: fig-color_diversity
#| fig-cap: Unique colour count by year
#| echo: false

# Creating dataframe for yearly_summary
yearly_summary = sets |>
  select(set_num, year, theme_id) |>
  inner_join(inventories, by = "set_num") |>
  inner_join(inventory_parts, by = c("id" = "inventory_id")) |>
  filter(color_id >= 0) |>
  distinct(set_num, year, color_id, theme_id) |>
  group_by(year) |>
  summarize(n_sets = n_distinct(set_num),
            unique_colors = n_distinct(color_id),
            n_themes = n_distinct(theme_id),
            .groups = "drop") |>
  filter(!is.na(year))

# Plot unique colors over years
yearly_summary_plot = ggplot(yearly_summary, aes(x = year, y = unique_colors)) +
  geom_area(fill = "#00BC8C", alpha = 0.2) +
  geom_line(color = "#00BC8C", linewidth = 1) +
  geom_point(color = "#00648C") +
  geom_smooth(method = "loess", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Number of unique colors used in lego sets over the years",
       x = "Year",
       y = "Unique Colors") +
  ylim(0,125) +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
#| label: fig-most_pieces
#| fig-cap: Top 20 sets by part count with color proportion
#| echo: false

# Creating a hex column to combine # with the hex rgb values in colors$rgb
colors = colors|>
  mutate(hex = paste0("#", rgb))

# Summarising parts by set and colour
set_parts_colors = inventory_parts |>
  inner_join(inventories, by = c("inventory_id" = "id")) |>
  group_by(set_num, color_id) |>
  summarise(parts_color = sum(quantity, na.rm = TRUE), .groups = "drop")

# Finding the total parts per set
set_total_parts = set_parts_colors |>
  group_by(set_num) |>
  summarise(total_parts = sum(parts_color), .groups = "drop")

# Joining the dataframes together
set_parts_colors = set_parts_colors |>
  inner_join(set_total_parts, by = "set_num") |>
  inner_join(sets |> select(set_num, name), by = "set_num")

# Calculating proportion of color for each set
set_parts_colors = set_parts_colors |>
  mutate(proportion = parts_color/total_parts)

# Retrieving the top 20 sets by total parts
top_20_sets = set_total_parts |>
  arrange(desc(total_parts)) |>
  slice_head(n = 20) |>
  pull(set_num)

# Joining with the top 20 sets with hex color
top_20_data = set_parts_colors |>
  filter(set_num %in% top_20_sets) |>
  left_join(colors |> select(id, hex), by = c("color_id" = "id"))

# Dropping unused datasets to improve performance
rm(top_20_sets)

# Plotting the top 20 sets with the most piece and filled with proportion of colours
top_20_plot = ggplot(top_20_data, 
                     aes(x = reorder(name, total_parts), 
                         y = parts_color, fill = hex)) +
  geom_bar(stat = "identity", 
           color = "black", 
           linewidth = 0.1) +
  scale_fill_identity() +
  labs(title = "Top 20 lego sets by # parts and colour proportion",
       x = "",
       y = "Number of Parts",
       fill = "Color") +
  coord_flip() +
  theme_minimal() +
  theme(plot.title.position = "plot",plot.title = element_text(hjust = 0))

# <--Remove # to clean all previous selections 
# rm(colors, least_popular, most_popular, part_year, parts, yearly_summary, theme_name_summary, theme_correlations, sets, parts, inventory_parts, set_parts_colors, set_themes, set_total_parts, theme_correlation, themes, inventories, top_20_data)
```

### LEGO's evolution: Insights into theme popularity and set complexity

<br><br>

#### Group Members:

1.  William Tang
2.  Prem MV
3.  Rong Chen
4.  Zhaoyu Zhang

------------------------------------------------------------------------

#### Data-set Overview

-   **Source:** TidyTuesday 2022 LEGO dataset - Rebrickable.

-   **Description:**

    -   Themes & Colours\
    -   Sets & Release year\
    -   Number of parts per set\
    -   Inventory & Parts

-   **Scope:** No sales or revenue data—only set metadata.

------------------------------------------------------------------------

#### Analysis Questions

1.  What were the most popular and least popular themes in LEGO?
2.  How has the complexity of LEGO sets changed over the years?
3.  What are the largest LEGO sets available officially?

------------------------------------------------------------------------

::::: columns
#### Methodology

::: {.column width="50%" style="font-size: 0.7em;"}
-   Used TidyTuesday LEGO dataset from Rebrickable
-   Filtered sets with missing or zero part counts
-   Initial exploration: number of sets by theme and median part count
:::

::: {.column width="50%" style="font-size: 0.55em;"}
```{r}
#| echo: false
#| results: asis
#| message: false
#| warning: false

knitr::kable(
  head(theme_name_summary, 20),
  caption = "Theme ordered by sets with their median parts",
  row.names = FALSE
)
```
:::
:::::

------------------------------------------------------------------------

::::: columns
#### Theme popularity

::: {.column width="50%" style="font-size: 0.7em;"}
-   Grouped sets by theme and calculated set count and median part count
-   Top 20 most and least popular themes
-   Bar length = \# of sets, fill = median parts count
:::

::: {.column width="50%"}
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

most_popular_plot
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

least_popular_plot
```
:::
:::::

------------------------------------------------------------------------

::::: columns
#### Trend over time

::: {.column width="50%" style="font-size: 0.7em;"}
-   Analysed median part counts and colour diversity by year
-   Added trend lines for better visibility
-   Focused on identifying how LEGO sets evolved over time
:::

::: {.column width="50%"}
```{r}
#| label: trend
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

part_year_plot
```

```{r}
#| label: diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

yearly_summary_plot
```
:::
:::::

------------------------------------------------------------------------

::::: columns
#### Largest Sets Analysis

::: {.column width="50%" style="font-size: 0.7em;"}
-   Top 20 largest LEGO sets by part count
-   Bars segmented by actual hex colours to show colour usage per set
-   Highlights set complexity and visual diversity
:::

::: {.column width="50%"}
```{r}
#| label: most_pieces
#| echo: false
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 7

top_20_plot
```
:::
:::::

------------------------------------------------------------------------

::::: columns
#### Results: Most popular and least popular LEGO themes

::: {.column width="50%" style="font-size: 0.7em;"}
**Most popular**

-   Strong brand identity\
-   Wide audience appeal\
-   Ongoing product development

**Least popular**

-   Short lived\
-   Targeted at small audiences\
-   Limited by licensing
:::

::: {.column width="50%"}
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

most_popular_plot
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

least_popular_plot
```
:::
:::::

------------------------------------------------------------------------

::::: columns
#### Results: Complexity of LEGO sets over the years

::: {.column width="50%" style="font-size: 0.7em;"}
-   LEGO sets became more complex up to the late 1970s
-   Median complexity has plateaued since then
-   Diversification of set sizes to meet the needs of different age groups and skill levels
-   strong indicator of increased design sophistication and build complexity
:::

::: {.column width="50%"}
```{r}
#| label: trend_explanation
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

part_year_plot
```

```{r}
#| label: diversity_explanation
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5.5
#| fig-height: 3.3

yearly_summary_plot
```
:::
:::::

------------------------------------------------------------------------

::::: columns
#### Results: Sets with the most parts and color proportions

::: {.column width="50%" style="font-size: 0.7em;"}
-   The number of parts in a LEGO set often correlates with either extensive design detail or technical function
-   Some sets aim for vibrant storytelling vs precision and realism
:::

::: {.column width="50%"}
```{r}
#| label: most_pieces_explanation
#| echo: false
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 7

top_20_plot
```
:::
:::::

------------------------------------------------------------------------

#### Results - Extra Discussion

More popular themes have on average significantly fewer parts in comparison to the least popular themes

-   Accessibility
-   Simplicity
-   Recognisability
-   Rather than set complexity alone

Simpler sets can appeal to more customers, making them more commercially successful

------------------------------------------------------------------------

#### Results - Extra Discussion

Correlation analysis returned -0.08

-   Very weak negative correlation
-   Little to no linear relationship between a theme's popularity and complexity of its sets

Finding is aimed to emphasize simplicity and accessibility are important factors of Lego theme popularity

------------------------------------------------------------------------

#### Conclusion

-   There are themes that LEGO focuses more on producing
-   LEGO is definitely getting more complex as the years go on
-   LEGO's most successful themes strike balance between accessibility and innovation, highlighting the importance of thoughtful design over sheer complexity

------------------------------------------------------------------------

## 
